---
title: "AQM_Nov9_Advertising"
author: "David Dvorak"
date: "November 12, 2016"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Due Nov. 16, 2016

Derive and manually compute the following in R, then cross-check them to the R output from `lm()` (use Advertising.csv, as used in class):

* Estimates of the coefficients using matrix notation.
* Residuals using matrix notation.
* Plot residuals against each explanitory variable. Interpret.

```{r load_packages, message=FALSE, results='hide',include=FALSE}
library(ggplot2)
#setwd("~/GitHub/Dvorak_David/Advertising_and_Regression")
```
Input the data on Sales vs. spending on TV, Newpaper and Radio advertising
```{r}
dat.adv <- read.csv("Advertising.csv")
Sales <- dat.adv$Sales
TV <- dat.adv$TV
Radio <- dat.adv$Radio
Newspaper <- dat.adv$Newspaper
```
Linear least squares regression for each of the possible explanatory variables
```{r}
mod <- lm(Sales ~ TV+Radio+Newspaper, dat.adv)
```
Residuals for linear least squares fitting
```{r}
res <- residuals(mod)
```

Plot Sales vs. Explanatory Variables
```{r}
qplot(TV,Sales) # tilda usage: Sales "depends on" TV. Equivalent to plot(TV,S)
qplot(Radio,Sales)
qplot(Newspaper,Sales)
```
